On the eletrodynamic of moving bodies

### 2019-05-16 Article_AS.Rmd
Materiel et methodes

@@

# Introduction
Thousands of hectares of Canadian forest are harvested each year in addition to losses due to insects, diseases and fires. It is therefore essential to offset these losses if we maintain Canada's forests. Genetic improvement, which is a tool for reforestation, is becoming a crucial and increasingly necessary discipline for the development of the Canadian forest. Since trees are organisms that live for several decades, it is essential to take into account the current climate change context by using trees capable of adapting to current and future climatic conditions. It is in this context that the genetic improvement intervenes so that the reforestation is made from tree to adapt to the climate and having interesting characters for wood production. Improvement programs are in place for the most important commercial species, including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions (Canadian Council of Forest Ministers 2009). \

@@

>>>>

<!-- JR: Commentaire -->

@@

# Introduction

Thousands of hectares of Canadian forest are harvested each year in addition to losses due to insects, diseases and fires. It is therefore essential to offset these losses if we maintain Canada's forests. Genetic improvement, which is a tool for reforestation, is becoming a crucial and increasingly necessary discipline for the development of the Canadian forest. Since trees are organisms that live for several decades, it is essential to take into account the current climate change context by using trees capable of adapting to current and future climatic conditions. It is in this context that the genetic improvement intervenes so that the reforestation is made from tree to adapt to the climate and having interesting characters for wood production. Improvement programs are in place for the most important commercial species, including white spruce (Picea glauca (Moench) Voss) which has a significant economic value because of its geographical distribution and abundance. Spruce wood is valuable as lumber, material for carpentry products and for pulping because of the characteristics of its fibres [@Zhang2008]. In addition, the plasticity of spruce, enable this species to adapt to various types of soil and climatic conditions [@Nienstaedt1990]. All these characteristics make white spruce a species of great interest in North America and consequently it is heavily reforested in various silvicultural systems in Canada with more than 150 million of natural an hybrid seedlings planted yearly accross almost all forested regions which represents a wide range of soils types and a climatic conditions (Canadian Council of Forest Ministers 2009). \

@@

## Sampling
The dataset used in this study came from genetic selection for a series of tests according to provenances [@Beaulieu1996]. The improvement program for this region was led by Quebec’s Ministry of Forests, Wildlife and Parcs, Canada. Experimental plantings were initiated from white spruce seedlings grown in nurseries and planted at two years of age at a regular spacing between trees of 2 m in all directions. This series of tests was established in 1999 by the Canadian Forest Service and is one of the three essential parts of the white spruce breeding program in Quebec. Two sites were selected for sampling: one near Saint-Casimir, Quebec (46.6994, -72.1119), in the maple–basswood ecological zone (warm weather) , and the other one near Cabano (Asselin), Quebec (47.8808, -68.4436), in the balsam fir-yellow birch ecological zone (cool weather). Between nine and 14 trees per site were sampled from each of the 93 full-sib families, for a total sample of 2365 trees (Table 1). Trees were 15 years old at the time of sampling. Trees were assigned to row-plots of five trees/plot (2 m × 2 m spacing). Core samples were extracted at breast height (1.3 m above ground level) from the southern side of the trees to minimize the variation due to sample orientation. Tree height and diameter at breast height (DBH, 1.3 m above ground) were recorded in 2015 at the age of 15 years, and a wood increment core was extracted from the south facing side of each tree. Cores were stored in a freezer, conditioned to 7% moisture and cut to 1.68 mm thickness prior to X-ray densitometry analyses (Quintek Measurement Systems, TN). Wood density was calculated as a ring area weighted mean from recorded pith to bark wood density profiles.

Study sites (image)

Growth ring density and ring width measurements were taken from pith-to-bark, using an X-ray densitometer (Quintek Measurement Systems, TN) at a resolution of 0.0254 mm, and the data were reported as relative wood density on a dry weight basis. To verify the accuracy of dates and determine if there were missing rings, we used COFECHA, which is a dendrochronology software program used for cross-dating and quality control of annual ring measurements [@HOLMES1983].

## Model Development
A model was developed to describe the patterns of variation in mean annual ring density in the stem in white spruce. 
Considering the hierarchical structure of the data, with annual rings nested in the cores, cores nested in the trees, and trees nested in blocks, we opted for a mixed-effects modelling approach using nlme package [@NLME2018] in the R statistical programming environment [@R2018]. This type of model also considers the autocorrelation between the observations [@Lindstrom1990]. Thus, the estimates of the parameters can vary randomly around the average values of each level of the data structure (site, block, tree, core) [@Auty2013]. As in the study of @Xiang2014, pith-to-bark wood density profiles in our data were different among cores. A nonlinear equation was fit for each core to describe ring density profiles, producing different parameter estimates for each core. The radial pattern of wood density in white spruce can be described as follows:

```
RD = b0*e ( b1/CA ) + CA/b2 + e

```
where RD is the mean ring density (kg∙m-3), CA is cambial age (years), b0, b1 and b2 are empirically determined parameters, and e is the residual error term. Parameter b0 denotes the intercept when CA approaches a theoretical value of 0. We noticed that it was impossible to optimize the equation using the same estimated parameters for all tree cores, because the pattern of variation of annual rings average density of each tree is more or less different from the others. We, therefore, used the optim function of the stats package in the R [@R2018] statistical programming environment to find the best parameters iteratively for each tree core . The fit of the model was evaluated by visual analysis of the model residual curves against adjusted values and explanatory variables, fit indices (R2) (Parresol 1999). The graphic description in figure 1 of the three parameters of our model makes it possible to
